---
title: "MachineLearning_assignment"
author: "biohuhu"
date: "Sunday, July 26, 2015"
output: html_document
---

### Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).   

### Methods
Use the caret package:
```{r}
library(caret)
```
Read in  training and testing data
```{r}
testing= read.csv("pml-testing.csv", header = TRUE, row.names = 1, fill = TRUE)
training = read.csv("pml-training.csv", header = TRUE, row.names = 1, fill = TRUE)
head(training)
head(testing)
```
There are many missing values. Have to remove the columns that have missing values
```{r}
training2 = training[, colSums(is.na(training) | training=="")==0]
```
remove the first 10 columns which are not predictors that can be expanded to new data sets
```{r}
training3 = training2[,-10:-1]
```
   
Now partition the training data further into test and training data to estimate prediction accuracy
```{r}
trainIndex = createDataPartition(training3$classe, p = 0.75, list=F)
training4 = training3[trainIndex,]
testing4 = training3[-trainIndex,]

modelFit = train(classe~., method = "rpart", data = training4)
#print(modelFit$finalModel)
prediction=predict(modelFit, newdata = testing4)
confusionMatrix(testing4$classe, prediction)
```
0.49 is the accuracy  

I attempted to use boosting (method = "gbm"") and random forrest (method = "rf"), but my computer doesn't have enough memory to handle the computation.  

To tune parameters, we can use a 5 fold Cross Validation using training data
```{r}
# define training control
train_control <- trainControl(method="cv", number=5)
# train the model 
model <- train(classe~., data=training3, trControl=train_control, method="rpart")
model
```
The best model out of the 5 fold CV is 0.50.   
Use it to make predictions for testing data, and write into files:
```{r}
prediction=predict(model, newdata = testing)
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(as.character(prediction))
```
The files are uploaded for grading. The overall accuracy is 40%, which is lower than the 5 fold CV accuracy (50%). Out of sample accuracy being lower is expected.



